\markboth{Acknowledgments}{}
%%\engExpl{It is nice to acknowledge the people that have helped you. It is
%%  also necessary to acknowledge any special permissions that you have gotten –
%%  for example, getting permission from the copyright owner to reproduce a
%%  figure. In this case, you should acknowledge them and this permission here
%%  and in the figure’s caption. \\
%%  Note: If you do \textbf{not} have the copyright owner’s permission, then you \textbf{cannot} use any copyrighted figures/tables/\ldots . Unless stated otherwise all figures/tables/\ldots are generally copyrighted.
%%}

\begin{quote}
    “ Persistence and resilience only come from having been given the chance to work through difficult problems „

    -- Gever Tulley
\end{quote}
The work that brought me to complete this master thesis made me grow incredibly, laying a foundation of the person I am today. This would have not been possible without other individuals, who I would like to thank in this section.

I would like to thank Jim Dowling for giving me the opportunity to work on this project and Salman Niazi that helped me make it happen. Many thanks to also all the Hopsworks AB employees, who welcomed me from day one.

I would like to thank my KTH examiner Vladimir Vlassov, and my two supervisors Fabian Schmidt and Sina Sheikholeslami for their many corrections to my numerous drafts.

I would like to thank all my friends that in the last years made me feel part of a larger family. Especially my Milan group, to which I will never be tired to come back to: Sebastiano, Virginia, Luca, Alfonso, Giacomo and Andrea. A special mention goes to Sebastiano that in the last year was always by my side, an amazing partner in an ever changing world. Even if now we parted ways, I would still love to visit you in any city you might be located in.

I would like to thank my parents and my brothers. In my hardest moment of difficulty they were able to help me get back on track to complete this project, while supporting me all the way. You enabled me to perform my best now and to do so in the future.

Finally I would like to thank my loved one, Elena who has been by my side well before I started my academic path. You were my space of serenity even in the darkest hour. You brought me joy when I felt I couldn't feel any. I look forward to our present and our future together.
