% The first abstract should be in the language of the thesis.
  \markboth{\abstractname}{}
\begin{scontents}[store-env=lang]
eng
\end{scontents}
%%% The contents of the abstract (between the begin and end of scontents) will be saved in LaTeX format
%%% and output on the page(s) at the end of the thesis with information for DiVA facilitating the correct
%%% entry of the meta data for your thesis.
%%% These page(s) will be removed before the thesis is inserted into DiVA.
%% \engExpl{All theses at KTH are \textbf{required} to have an abstract in both \textit{English} and \textit{Swedish}.}
%% \engExpl{Exchange students may want to include one or more abstracts in the language(s) used in their home institutions to avoid the need to write another thesis when returning to their home institution.}
%% \generalExpl{Keep in mind that most of your potential readers are only going to read your \texttt{title} and \texttt{abstract}. This is why the abstract must give them enough information so that they can decide if this document is relevant to them or not. Otherwise, the likely default choice is to ignore the rest of your document.\\
%% An abstract should stand on its own, i.e., no citations, cross-references to the body of the document, acronyms must be spelled out, \ldots .\\Write this early and revise as necessary. This will help keep you focused on what you are trying to do.}
\begin{scontents}[store-env=abstracts,print-env=true]
The need to build Machine Learning (ML) models based on ever-increasing amounts of data brought new challenges to data management systems. Feature stores have emerged as a centralized data platform enabling feature reuse while organizing data transformations and ensuring consistency between feature engineering, model training, and inference. Recent publications demonstrate that the Hopsworks feature store outperforms existing cloud-based alternatives in training and online inference query workloads. In its offline feature store, the Hopsworks feature store stores batch or historical data, collecting it into feature groups, i.e., logical tables of figures, organized in Apache Hudi tables, i.e., a data management layer, and stored on HopsFS, Hopsworks HDFS distribution. However, even in this system, the latency to perform a write operation is at least one or more minutes, even for small quantities of data (1 GB or less). A previous improvement in read latency using an Arrow Flight and DuckDB server suggests that this limitation is caused by Spark, which the system uses to write data on Apache Hudi tables. 
A promising approach to avoid using Spark appears to be adopting Delta Lake over Apache Hudi to manage and access data using a Rust library called delta-rs. This thesis investigates the possibility of reducing the read and write latency in the offline feature store by expanding the delta-rs library to support the Hopsworks feature store file system called HopsFS and comparatively evaluating the performance of the legacy and newly implemented system. Two major iterations of storage support in delta-rs for HopsFS were developed to meet the strict production-ready requirements defined before development. The system was then evaluated by performing and measuring read and write operations in four different CPU configurations, increasing the number of CPU cores up to eight. Experiments were performed fifty times to estimate a confidence interval, allowing an accurate comparative evaluation of the systems. Results confirmed the superior performance of the delta-rs library over the Spark system in all write operations with a latency reduction from ten up to forty times. Delta-rs also surpassed the Spark-alternative in read operations with a latency reduction of forty-seven percent, up to forty times. These findings encourage future research investigating Spark-alternative when optimizing performance in small-scale (1 GB - 100 GB) data management systems. The system developed will find application in the Hopsworks feature store production environment.
\end{scontents}
\begin{comment}
\engExpl{The following are some notes about what can be included (in terms of LaTeX) in your abstract.}
Choice of typeface with \textbackslash textit, \textbackslash textbf, and \textbackslash texttt:  \textit{x}, \textbf{x}, and \texttt{x}.

Text superscripts and subscripts with \textbackslash textsubscript and \textbackslash textsuperscript: A\textsubscript{x} and A\textsuperscript{x}.

Some symbols that you might find useful are available, such as: \textbackslash textregistered, \textbackslash texttrademark, and \textbackslash textcopyright. For example, 
the copyright symbol: \textbackslash textcopyright Maguire 2022 results in \textcopyright Maguire 2022. Additionally, here are some examples of text superscripts (which can be combined with some symbols): \textbackslash textsuperscript\{99m\}Tc, A\textbackslash textsuperscript\{*\}, A\textbackslash textsuperscript\{\textbackslash textregistered\}, and A\textbackslash texttrademark resulting in \textsuperscript{99m}Tc, A\textsuperscript{*}, A\textsuperscript{\textregistered}, and A\texttrademark. Two examples of subscripts are: H\textbackslash textsubscript\{2\}O and CO\textbackslash textsubscript\{2\} which produce  H\textsubscript{2}O and CO\textsubscript{2}.

You can use simple environments with begin and end: itemize and enumerate and within these use instances of \textbackslash item.

The following commands can be used: \textbackslash eg, \textbackslash Eg, \textbackslash ie, \textbackslash Ie, \textbackslash etc, and \textbackslash etal: \eg, \Eg, \ie, \Ie, \etc, and \etal.

The following commands for numbering with lowercase Roman numerals: \textbackslash first, \textbackslash Second, \textbackslash third, \textbackslash fourth, \textbackslash fifth, \textbackslash sixth, \textbackslash seventh, and \textbackslash eighth: \first, \Second, \third, \fourth, \fifth, \sixth, \seventh, and \eighth. Note that the second case is set with a capital 'S' to avoid conflicts with the use of second of as a unit in the \texttt{siunitx} package.

Equations using \textbackslash( xxxx \textbackslash) or \textbackslash[ xxxx \textbackslash] can be used in the abstract. For example: \( (C_5O_2H_8)_n \)
or \[ \int_{a}^{b} x^2 \,dx \]
Note that you \textbf{cannot} use an equation between dollar signs.

Even LaTeX comments can be handled, for example: \% comment.
Note that one can include percentages, such as: 51\% or \SI{51}{\percent}.
\end{comment}
\subsection*{Keywords}
\begin{scontents}[store-env=keywords,print-env=true]
% If you set the EnglishKeywords earlier, you can retrieve them with:
\InsertKeywords{english}
% If you did not set the EnglishKeywords earlier then simply enter the keywords here:
%First keyword, Second keyword, Third keyword, Fourth keyword
\end{scontents}
%%\engExpl{\textbf{Choosing good keywords can help others to locate your paper, thesis, dissertation, \ldots and related work.}}
%%Choose the most specific keyword from those used in your domain, see for example: the ACM Computing Classification System ({\small \url{https://www.acm.org/publications/computing-classification-system/how-to-use})},
%%the IEEE Taxonomy ({\small \url{https://www.ieee.org/publications/services/thesaurus-thank-you.html}}), PhySH (Physics Subject Headings)\linebreak[4] ({\small \url{https://physh.aps.org/}}), \ldots or keyword selection tools such as the  National Library of Medicine's Medical Subject Headings (MeSH)  ({\small \url{https://www.nlm.nih.gov/mesh/authors.html}}) or Google's Keyword Tool ({\small \url{https://keywordtool.io/}})\\

%%\textbf{Formatting the keywords}:
%%\begin{itemize}
%%  \item The first letter of a keyword should be set with a capital letter, and proper names should be capitalized as usual.
%%  \item Spell out acronyms and abbreviations.
%%  \item Avoid "stop words" - as they generally carry little or no information.
%%  \item List your keywords separated by commas (",").
%%\end{itemize}    
%%Since you should have both English and Swedish keywords - you might think of ordering them in corresponding order (\ie, so that the n\textsuperscript{th} word in each list correspond) - this makes it easier to mechanically find matching keywords.